### 2007-01-19 Documentation/part2.tex
*** empty log message ***

\index{norm}{AEDL@\AEDL|(}
\part{The Abstract Experiment Definition Language -- AEDL}
\label{part:2}

@@

Here, the general concept of the AEDL is described and illustrated by an example. In addition, a short introduction to the syntax of the AEDL is given.

@@

The goal of AEDL is to describe a large number of complex and very 
different experiments by a limited number of parameters. It allows a
representation of very different setups within one data structure, and thus implements universal rate and $\chi^2$ computation methods. For experiment simulations, usually a new piece of code is written and compiled
for each different experiment. In many cases, even parameter changes, such as
the number of bins, require the recompilation of the source code. 
However, such a technique soon reaches its limits when the simulated experiments are rather complex, or more than one type of experiment is studied simultaneously. Furthermore, it is very difficult to verify the correctness of the obtained results, since every time a new piece of code is added to 
deal with a new experiment type, new errors will be introduced.

@@

In order to implement an experiment definition including various
sources of systematical errors, we use several abstraction levels. 
The first level is the so-called ``channel'', \index{norm}{Channel}
 which is the link between 
the oscillation physics and the detection properties for a specfific oscillation pattern (\cf, \figu{channel}). A channel specifies the mapping of a specific neutrino flavor produced by the source onto a reconstructed neutrino flavor.
For example, a muon neutrino oscillates into an electron neutrino and subsequently interacts via quasi-elastic charged current scattering. The measured energy and direction of
the secondary electron in the detector then allows to reconstruct the neutrino energy. The connection from the source flux of the muon neutrino, via the  probability to appear as a electron neutrino, to its detection properties (such as cross sections and energy smearing) is encapsulated into the channel.

@@

The channels are the building blocks for the so-called ``rules''.
\index{norm}{Rule} In general, a rule consists of one or more ``signal'' and ``background'' oscillation channels, which are normalized with efficiencies
(\cf, \figu{rule}). The event numbers from these channels are added {\em before} the $\Delta \chi^2$-value is calculated.\footnote{Note that
in this manual, the $\chi^2$ and $\Delta \chi^2$ are equal, since
for simulated data $\Delta \chi^2 = 0$ at the best-fit point. Thus, we
are using $\chi^2$ and $\Delta \chi^2$ as equal quantities.} In addition, each rule implements {\em independent} systematics, such as signal and background normalization errors. Eventually, each rule gives a $\Delta \chi^2$-value, and the total $\Delta \chi^2$ of one experiment is obtained by adding the $\Delta \chi^2$'s of all rules (\cf, \figu{experiment}). 
 An example for a rule could look like this: We want to detect electron
  neutrino appearance (``signal''), where the overall efficiency for 
  quasi-elastics electron neutrino events is $0.4$. There is a fraction of 
  $0.01$ of all neutral current events which are mis-identified as 
  quasi-elastic electron neutrino events (``background''). The neutral 
  current fraction is only known within $10\%$ (``background uncertainty'') 
  and there is
an energy scale uncertainty of $100\,\mathrm{MeV}$ (``energy calibration error'').
All this systematics is independent of the other rules.  Thus, a rule connects the event rates to the calculation of a $\Delta \chi^2$ which properly includes systematical errors. The resulting $\Delta \chi^2$ is then the starting point for the oscillation physics analysis. Note again that
\begin{itemize}
\item
 Within each rule, the event numbers are added.
\item
 Within each rule, the systematics is treated independently from the other rules.
\item
 For each rule, the $\Delta \chi^2$ is computed; the $\Delta \chi^2$'s from all rules are added.
\end{itemize}

@@

\section{A simple example for AEDL}

Experiments in \GLOBES\ are defined by the Abstract Experiment Definition Language (AEDL). The experiment definition is written into a text file using the AEDL syntax. Currently, a number of pre-defined experiment definition files are provided with \GLOBES , which have to be modified manually in order to define new experiments.  The application software then uses this text file to initialize the experiment, where other secondary files might be read for source fluxes, cross sections \etc . In this section, we show the definition of a very simple neutrino factory in AEDL, where we do not go into details. In the next chapter, we will discuss each of the individual steps in detail.

@@

The experiment defined here represents a first simplified version of a neutrino factory experiment. It still lacks the correct energy dependence of the efficiencies, the antineutrino disappearance channel, and the channels and rules for the symmetric operation with $\mu^-$ stored. However, it may serve as a simple, introductory example. In the next chapter, we will demonstrate that AEDL is much more powerful than illustrated here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction to the syntax of AEDL}
\label{sec:syntax}

We now give a short introduction to the syntax of AEDL.
 The first eight characters have to be {\tt \%!GLoBES}
in order to avoid parsing megabytes of chunk
 and producing thousands of error messages.  In addition, the minimum
\GLOBES\ version that the \AEDL\ file is supposed to run with has to be
defined by a {\tt \$version} statement, such as
{\tt \$version="3.0.0"}.
\index{aedl}{version@{\tt \$version}}
%
Comments can be used in the same way as in C:
\begin{quote}
{\tt /* This starts a comment\\
 and here the comment ends */ \\
// Another comment
}
\end{quote}
There are pre-defined variables which all start with {\tt \$}. Their range
is also checked. For example,  {\tt 
\$bins} can be only between $0$ and $500$.\footnote{The upper limit is 
only there for safety reasons, the memory is allocated dynamically.} If one uses a {\tt float} quantity where  an {\tt int} is expected, the {\tt float} will be converted to an {\tt int} in the same way as in C.  For example, we have scalar variables
\begin{quote}
{\tt
\$bins = 10\\
\$baseline = 1200.0
}
\end{quote}
and simple lists
\begin{quote}
{\tt
\$densitytab=\{1.0,2.2343,3.3432\} 
}
\end{quote}
%
Since there are often groups of data which we want to refer to later,
environments can be used. This is illustrated 
with the channel definition part:
\begin{quote}
{\tt channel(\#ch1)<\\
\tb  $\ldots$\\
>
}
\end{quote}
The first part is the type of environment, which is {\tt channel} here. 
There are the following types of environments in AEDL:
\begin{quote}
{\tt nuflux\\
cross\\
channel\\
energy\\
rule
}
\end{quote}
Besides the environment type, there is a user-defined name 
beginning with {\tt \#}
in the above example: {\tt \#ch1}. It can be used later to refer to the 
channel defined in {\tt <$\ldots$>}. Those names are so-called 
``automatic variables'' and have to start with {\tt \#}. Note that these names have to be unique and can only be refered to after their definition.
However, similar to C, one can give a declaration without definition before:
\begin{quote}
{\tt    channel(\#ch2)<>}
\end{quote}
Now one can refer to the name {\tt \#ch2}, while the actual channel definition comes later. The internal representation of this automatic
variable is a number, which obtains its value from a counter for each type of environment. For example, for {\tt channel} the counter is {\tt numofchannels}. The counter keeps track of how many different names 
there are for one type of environment, which means that it counts the number of channels, rules, energy resolution functions, \etc . Thus, the automatic
variables are numbered in the order of their definition, and the number
can later be used to refer to them in the C code (from $0$ to {\tt numof...}$-1$). In order to facilitate the the mapping from names in AEDL to indices
in C there are two functions \GLB{NameToValue} and \GLB{ValueToName} which
make this transition (see \Sec~\ref{sec:aedl_names}, 
page~\pageref{sec:aedl_names}).

@@

\index{aedl}{include@{\tt include}}
As another feature of AEDL one, can use include files with the {\tt include} command. Includes can be nested up to {\tt MAX\_INCLUSION\_DEPTH}, which is currently set to $10$. Error reporting works 
 for nested includes, too. The included file is not required to begin 
 with {\tt \%!GLoBES} to facilitate cut and paste:
\begin{quote}
{\tt include "./file\_1"}
\end{quote}
With this include mechanism, one can use constructions such as 
\begin{quote}
{\tt    include  "Exp1.glb"\\
        \#NEXT\#\\
        include   "Exp2.glb"
}
\end{quote}
in order to initialize a combined analysis of the experiments defined in the files {\tt Exp1.glb} and {\tt Exp2.glb}. Note that one has 
to use quotation marks for filenames in \AEDL.\index{norm}{File names}
Even if one uses the
automatic variable {\tt \#CC} in both experiments, 
but the cross section data are different (for example, because of different target nuclei), the correct 
cross section data will be applied to each of the experiments. 
Note that, alternatively, one can 
also load both files successively by two separate calls of 
\GLB{InitExperiment}.

@@

In addition, some simple algebraic manipulations are possible, such as
\begin{quote}
{\tt
Pi+1\\
\verb+Pi^2+\\
sin(Pi/2)
}
\end{quote}
The following mathematical functions from {\tt <math.h>} are available: 
{\tt sin}, {\tt cos}, {\tt tan}, {\tt asin}, {\tt acos}, {\tt atan}, 
{\tt log}, {\tt log10}, {\tt exp}, {\tt sqrt}.
\index{aedl}{sin@{\tt sin}}%
\index{aedl}{cos@{\tt cos}}%
\index{aedl}{tan@{\tt tan}}%
\index{aedl}{asin@{\tt asin}}%
\index{aedl}{acos@{\tt acos}}%
\index{aedl}{atan@{\tt atan}}%
\index{aedl}{log@{\tt log}}%
\index{aedl}{log10@{\tt log10}}%
\index{aedl}{exp@{\tt exp}}%
\index{aedl}{sqrt@{\tt sqrt}}%
% 
These functions can be used everywhere, where
otherwise only a scalar number would appear. However, they can not be
applied to lists, \ie, expressions like {\tt sin(\{1,2,3\})} will not work.

@@

\section{More advanced AEDL features}
\label{sec:advaedl}

In \GLOBES\ 3.0 and higher, a number of new features can be used. The
features in this section are slighlty more advanced and can be skipped in a first reading of
the manual. The
most important one are lists as variables in \AEDL . They start with {\tt \%}, such as
\begin{quote}
{\tt
\%effs = \{ 0.2, 0.4, 0.6, 0.8, 1.0, 1.0, 1.0 \}
}
\end{quote}
Functions can be threaded over lists, \ie, they will be applied to each element
of list, and return a list. Note that the original list will be destroyed by this
process. Therefore, it is necessary to create a copy of your list if you want to
use the original and the threading result. For that purpose, the {\tt copy}
funcion is provided:
\begin{quote}
{\tt
 listb = copy(lista) \\
 listb := lista // Alternative method 
}
\end{quote}
\index{aedl}{copy@{\tt copy}}
You will also need to use {\tt copy} when you assign a list to an experiment structure (see below).

Two helper functions {\tt bincenter()} 
\index{aedl}{bincenter@{\tt bincenter}} and {\tt samplingbincenter} \index{aedl}{samplingbincenter@{\tt samplingbincenter}} return lists with the central energies of the bins or sampling points, respectively. For example,
\begin{quote}
{\tt
 \%bc=bincenter() 
}
\end{quote}
A very useful new features is an interpolation function which can directly interpolate a number of
points and evaluate them at a different set of places. For example,
\begin{quote}
{\tt
 \%energ = \{ 4.0,20.0,50.0 \} \\
 \%effs = \{ 0.0,1.0,1.0 \} \\
 \%ires = interpolation(\%energ,\%effs,1,\%bc) 
}
\end{quote}
\index{aedl}{interpolation@{\tt interpolation}}
interpolates the points with x-values {\tt \%energ} and y-values {\tt \%effs} with the interpolation order one (linear interpolation, third parameter) and evaluates the interpolation result at the bin centers obtained above, \ie, it returns a list of the y-values at the places specified by the last parameter. The only allowed interpolation orders are ``1'' (linear) and ``2'' (cubic splines). This example creates
a neutrino factory energy threshold function linearly climbing from 0 to 1 between 4~GeV and 20~GeV. It can be directly used in a channel definition, \eg, 
\begin{quote}
{\tt
 channel(\#nu\_mu\_appearance)< \\
\hspace*{0.5cm}	@channel = \#mu\_plus: +: electron: muon: \#CC: \#MINOS\ \\
\hspace*{0.5cm}	@post\_smearing\_efficiencies = copy(\%ires) \\
>
}
\end{quote}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Experiment definition with AEDL}

@@

As we have disussed before, \GLOBES\ can only deal with point sources. Thus,  it is not possible to study effects of the finite size of the neutrino production region, such as in the sun or in reactor experiments with many
neutrino sources (\eg, KamLAND). Therefore, a neutrino source in \GLOBES\ can, in general, be characterized by the flux spectrum for each neutrino flavor, the CP sign (neutrinos or antineutrinos), and the total luminosity
of the source.

Before we come to the definition of the source properties, let us discuss
the total integrated luminosity of the experiment. In \GLOBES , the total number of events is in general proportional to the product of
\begin{equation}
\mathrm{Fid.~detector~mass}\,\left[\mathrm{kt/t}\right]\times 
\mathrm{Running~time} \,\left[\mathrm{yr}\right]\times\left\{ \begin{array}{c}
\mathrm{Source~power}\,\left[\mathrm{MW/GW}\right]\\
\mathrm{Useful~parent~decays}\,\left[\mathrm{yr}^{-1}\right]
\end{array}\right.\,.
\end{equation}
Thus, the source power corresponds to either the amount of energy produced per time frame in the target (such as for nuclear reactors or sources based on pion decay), or the useful parent particle decays per time frame (neutrino factories, beta beams). In addition, the definition of the source power makes only sense together with the flux normalization, the running time, and the fiducial detector mass in order to define the total integrated luminosity. Therefore, one can, in principle, use arbitrary units for these components as long as their product gives the wanted neutrino flux. However, it is
recommended to use normalizations such that the source power units are $\mathrm{MW}$ for a proton-based beam, and $\mathrm{GW}_\mathrm{thermal}$ for a reactor experiment. Correspondingly, the detector mass units should be kilotons for a proton-based beam, and tons for a reactor experiment. In any case it is a good
idea to document the choices made by the user by corresponding comments
in \AEDL. For more details on the luminosity implementation, see appendix on page~\pageref{app:flux}.

@@

\index{aedl}{nuflux@{\tt nuflux}!time@{\tt "@time}}
There are two principal ways to initialize a neutrino flux: 
Either one can use 
a built-in source, or one can provide a file. In both cases,
a flux is defined by the environment {\tt nuflux}, such as
\index{aedl}{nuflux@{\tt nuflux}}
\begin{quote}
  {\tt nuflux(\#name)<\\
\tb $\ldots$\\
\tb @time = 8.0 \\
>}
\end{quote}
with a running time of $8$ years. Note that the running time is used within
the {\tt nuflux} environment. This feature can be used to load the neutrino
and antineutrino fluxes in an accelerator experiment separately, in order
to combine them with different running times in the respective operation modes.
The name of the flux {\tt \#name} will later be referred to in the channel definitions. Note that \GLOBES\ versions older than 3.0 use the still supported {\tt flux} environment, which is different from {\tt nuflux} by an undocumented normalization factor 5.2 for user-defined fluxes. This difference is explained in the appendix on page~\pageref{app:flux}.

@@

Furthermore, the matter density profile along the baseline
has to be specified. The simplest profile is a constant matter density equal to the average matter density from the PREM~\cite{Stacey} onion shell model of the earth:\index{norm}{PREM| \see{Matter density}}
\index{norm}{Matter density!of the earth}
\begin{quote}
{\tt \$profiletype=1 }
\end{quote}
%
If you are using this option please cite reference~\cite{Stacey}\index{norm}{Referencing!matter profile data}.

@@

Cross sections\index{norm}{Cross section} will later be used as part of the 
channel definition (see \Sec~\ref{sec:channel}). Similar to the source 
fluxes, they are provided by the user as a data file:
\index{aedl}{cross@{\tt cross}}
\index{aedl}{cross@{\tt cross}!cross file@{\tt  "@cross\_file}}
\begin{quote}
{\tt cross(\#name)<}\\
{\tt \tb @cross\_file ="user\_file\_1.dat"}\\
{\tt >}
\end{quote}  
This cross section can later be refered to by {\tt \#name}.

Cross sections in \GLOBES\ are given as total cross section divided by energy:
\begin{equation}
\hat\sigma(E)=\sigma(E)/E\,\left[ 10^{-38}\,
\frac{\mathrm{cm}^2}{\mathrm{GeV}} \right]
\end{equation}
The software assumes that the cross section files are text files with 
seven columns and $1001$ lines of the form
\index{norm}{Cross section!file}
\index{aedl}{cross@{\tt cross}!cross file@{\tt  "@cross\_file}}
\begin{quotation}
$\mathrm{log}_{10} E\quad
\hat\sigma_{\nu_e}\quad
\hat\sigma_{\nu_\mu}\quad
\hat\sigma_{\nu_\tau}\quad
\hat\sigma_{\bar\nu_e}\quad
\hat\sigma_{\bar\nu_\mu}\quad
\hat\sigma_{\bar\nu_\tau}$
\end{quotation}
Here the logarithms of the energy values have to be equidistant. For 
arbitrary energies, linear interpolation is used. If the energy leaves the
range of values given in the file, $0.0$ will be assumed. In general, it is advisable to provide the cross sections in the range between {\tt \$sampling\_min} and {\tt \$sampling\_max} (\cf, \Sec~\ref{sec:energy}).
Cross sections for unused neutrino flavours have to be filled with zeros, and
can not just be omitted.

@@

In the following subsections we will explain the energy resolution function.
All energy resolution functions are defined within an {\tt energy} environment and can be refered to by {\tt \#name}:
\begin{quote}
  {\tt energy(\#name)<\\
\tb $\ldots$\\
>}
\end{quote}
The individual parameters of the environment will be defined below and depend on the algorithm used.

@@

\subsection{Low-pass filter}

@@

\subsection{Manual energy smearing}

@@

\subsection*{\AEDL\ external variable substitution}
\index{norm}{globes@{\tt globes}!variable substitution}

>>>>

\index{norm}{AEDL@\AEDL|(}
\part{The Abstract Experiment Definition Language -- \AEDL }
\label{part:2}

@@

Here, the general concept of AEDL\ is described and illustrated by an example. In addition, a short introduction to the syntax of \AEDL\ is given.

@@

The goal of \AEDL\ is to describe a large number of complex and very 
different experiments by a limited number of parameters. It allows a
representation of very different setups within one data structure, and thus implements universal rate and $\chi^2$ computation methods. For experiment simulations, usually a new piece of code is written and compiled
for each different experiment. In many cases, even parameter changes, such as
the number of bins, require the recompilation of the source code. 
However, such a technique soon reaches its limits when the simulated experiments are rather complex, or more than one type of experiment is studied simultaneously. Furthermore, it is very difficult to verify the correctness of the obtained results, since every time a new piece of code is added to 
deal with a new experiment type, new errors will be introduced.

@@

In order to implement an experiment definition including various
sources of systematical errors, we use several abstraction levels. 
The first level is the so-called ``channel'', \index{norm}{Channel}
 which is the link between 
the oscillation physics and the detection properties for a specific oscillation pattern (\cf, \figu{channel}). A channel specifies the mapping of a specific neutrino flavor produced by the source onto a reconstructed neutrino flavor.
For example, a muon neutrino oscillates into an electron neutrino and subsequently interacts via quasi-elastic charged current scattering. The measured energy and direction of
the secondary electron in the detector then allows to reconstruct the neutrino energy. The connection from the source flux of the muon neutrino, via the  probability to appear as a electron neutrino, to its detection properties (such as cross sections and energy smearing) is encapsulated into the channel.

@@

The channels are the building blocks for the so-called ``rules''.
\index{norm}{Rule} In general, a rule consists of one or more ``signal'' and ``background'' oscillation channels, which are normalized with efficiencies
(\cf, \figu{rule}). The event numbers from these channels are added {\em before} the $\Delta \chi^2$-value is calculated.\footnote{Note that
in this manual, the $\chi^2$ and $\Delta \chi^2$ are equal, since
for simulated data $\Delta \chi^2 = 0$ at the best-fit point. Thus, we
are using $\chi^2$ and $\Delta \chi^2$ as equal quantities.} In addition, each rule implements {\em independent} systematics, such as signal and background normalization errors. Eventually, each rule gives a $\Delta \chi^2$-value, and the total $\Delta \chi^2$ of one experiment is obtained by adding the $\Delta \chi^2$'s of all rules (\cf, \figu{experiment}). Note that each experiment shares a common matter density profile. 
 An example for a rule could look like this: We want to detect electron
  neutrino appearance (``signal''), where the overall efficiency for 
  quasi-elastics electron neutrino events is $0.4$. There is a fraction of 
  $0.01$ of all neutral current events which are mis-identified as 
  quasi-elastic electron neutrino events (``background''). The neutral 
  current fraction is only known within $10\%$ (``background uncertainty'') 
  and there is
an energy scale uncertainty of $100\,\mathrm{MeV}$ (``energy calibration error'').
All this systematics is independent of the other rules.  Thus, a rule connects the event rates to the calculation of a $\Delta \chi^2$ which properly includes systematical errors. The resulting $\Delta \chi^2$ is then the starting point for the oscillation physics analysis. Note again that
\begin{itemize}
\item
 Within each rule, the event numbers from different channels are added.
\item
 Within each rule, the systematics is treated independently from the other rules.
\item
 For each rule, the $\Delta \chi^2$ is computed; the $\Delta \chi^2$'s from all rules are added in an
experiment.
\end{itemize}

@@

\section{A simple example for \AEDL }

Experiments in \GLOBES\ are defined by the Abstract Experiment Definition Language (\AEDL ). The experiment definition is written into a text file using the \AEDL\ syntax. Currently, a number of pre-defined experiment definition files are provided with \GLOBES , which have to be modified manually in order to define new experiments.  The application software then uses this text file to initialize the experiment, where other secondary files might be read for source fluxes, cross sections \etc . In this section, we show the definition of a very simple neutrino factory in \AEDL , where we do not go into details. In the next chapter, we will discuss each of the individual steps in detail.

@@

The experiment defined here represents a first simplified version of a neutrino factory experiment. It still lacks the correct energy dependence of the efficiencies, the antineutrino disappearance rule, and the channels and rules for the symmetric operation with $\mu^-$ stored. However, it may serve as a simple, introductory example. In the next chapter, we will demonstrate that \AEDL\ is much more powerful than illustrated here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction to the syntax of \AEDL }
\label{sec:syntax}

We now give a short introduction to the syntax of \AEDL .
 The first eight characters have to be {\tt \%!GLoBES}
in order to avoid parsing megabytes of chunk
 and producing thousands of error messages.  In addition, the minimum
\GLOBES\ version that the \AEDL\ file is supposed to run with has to be
defined by a {\tt \$version} statement, such as
{\tt \$version="3.0.0"}.
\index{aedl}{version@{\tt \$version}}
%
Comments can be used in the same way as in C:
\begin{quote}
{\tt /* This starts a comment\\
 and here the comment ends */ \\
// Another comment
}
\end{quote}
There are pre-defined variables which all start with {\tt \$}. Their range
is also checked. For example,  {\tt 
\$bins} can be only between $0$ and $500$.\footnote{The upper limit is 
only there for safety reasons, the memory is allocated dynamically.} If one uses a {\tt float} quantity where  an {\tt int} is expected, the {\tt float} will be converted to an {\tt int} in the same way as in C.  For example, we have scalar variables
\begin{quote}
{\tt
\$bins = 10\\
\$baseline = 1200.0
}
\end{quote}
and simple lists
\begin{quote}
{\tt
\$densitytab=\{1.0,2.2343,3.3432\} 
}
\end{quote}
%
Since there are often groups of data which we want to refer to later,
environments can be used. This is illustrated 
with the channel definition part:
\begin{quote}
{\tt channel(\#ch1)<\\
\tb  $\ldots$\\
>
}
\end{quote}
The first part is the type of environment, which is {\tt channel} here. 
There are the following types of environments in \AEDL :
\begin{quote}
{\tt nuflux\\
cross\\
channel\\
energy\\
rule
}
\end{quote}
Besides the environment type, there is a user-defined name 
beginning with {\tt \#}
in the above example: {\tt \#ch1}. It can be used later to refer to the 
channel defined in {\tt <$\ldots$>}. Those names are so-called 
``automatic variables'' and have to start with {\tt \#}. Note that these names have to be unique and can only be referred to after their definition.
However, similar to C, one can give a declaration without definition before:
\begin{quote}
{\tt    channel(\#ch2)<>}
\end{quote}
Now one can refer to the name {\tt \#ch2}, while the actual channel definition comes later. The internal representation of this automatic
variable is a number, which obtains its value from a counter for each type of environment. For example, for {\tt channel} the counter is {\tt numofchannels}. The counter keeps track of how many different names 
there are for one type of environment, which means that it counts the number of channels, rules, energy resolution functions, \etc . Thus, the automatic
variables are numbered in the order of their definition, and the number
can later be used to refer to them in the C code (from $0$ to {\tt numof...}$-1$). In order to facilitate the the mapping from names in \AEDL\ to indices
in C there are two functions \GLB{NameToValue} and \GLB{ValueToName} which
make this transition (see \Sec~\ref{sec:aedl_names}, 
page~\pageref{sec:aedl_names}).

@@

\index{aedl}{include@{\tt include}}
As another feature of \AEDL , can use include files with the {\tt include} command. Includes can be nested up to {\tt MAX\_INCLUSION\_DEPTH}, which is currently set to $10$. Error reporting works 
 for nested includes, too. The included file is not required to begin 
 with {\tt \%!GLoBES} to facilitate cut and paste:
\begin{quote}
{\tt include "./file\_1"}
\end{quote}
With this include mechanism, one can use constructions such as 
\begin{quote}
{\tt    include  "Exp1.glb"\\
        \#NEXT\#\\
        include   "Exp2.glb"
}
\end{quote}
in order to initialize a combined analysis of the experiments defined in the files {\tt Exp1.glb} and {\tt Exp2.glb}. Note that one has 
to use quotation marks for filenames in \AEDL.\index{norm}{File names}
Even if one uses the
automatic variable {\tt \#CC} in both experiments, 
but the cross section data are different (for example, because of different target nuclei), the correct 
cross section data will be applied to each of the experiments. 
Note that, alternatively, one can 
also load both files successively by two separate calls of 
\GLB{InitExperiment}.

@@

In addition, some simple algebraic manipulations are possible, such as
\begin{quote}
{\tt
Pi+1\\
\verb+Pi^2+\\
sin(Pi/2)
}
\end{quote}
The following mathematical functions from {\tt <math.h>} are available: 
{\tt sin}, {\tt cos}, {\tt tan}, {\tt asin}, {\tt acos}, {\tt atan}, 
{\tt log}, {\tt log10}, {\tt exp}, {\tt sqrt}.
\index{aedl}{sin@{\tt sin}}%
\index{aedl}{cos@{\tt cos}}%
\index{aedl}{tan@{\tt tan}}%
\index{aedl}{asin@{\tt asin}}%
\index{aedl}{acos@{\tt acos}}%
\index{aedl}{atan@{\tt atan}}%
\index{aedl}{log@{\tt log}}%
\index{aedl}{log10@{\tt log10}}%
\index{aedl}{exp@{\tt exp}}%
\index{aedl}{sqrt@{\tt sqrt}}%
% 
These functions can be used everywhere, where
otherwise only a scalar number would appear.

@@

\section{More advanced \AEDL\ features$^*$}
\label{sec:advaedl}

In \GLOBES\ 3.0 and higher, a number of new features can be used.  The
most important one are lists as variables in \AEDL . They start with {\tt \%}, such as
\begin{quote}
{\tt
\%effs = \{ 0.2, 0.4, 0.6, 0.8, 1.0, 1.0, 1.0 \}
}
\end{quote}
Functions can be threaded over lists, \ie, they will be applied to each element
of list, and return a list. Note that the original list will be destroyed by this
process. Therefore, it is necessary to create a copy of your list if you want to
use the original and the threading result. For that purpose, the {\tt copy}
function is provided:
\begin{quote}
{\tt
 listb = copy(lista) \\
 listb := lista // Alternative method in environments
}
\end{quote}
\index{aedl}{copy@{\tt copy}}
You will also need to use {\tt copy} when you assign a list to an experiment structure (see below).

Two helper functions {\tt bincenter()} 
\index{aedl}{bincenter@{\tt bincenter}} and {\tt samplingbincenter()} \index{aedl}{samplingbincenter@{\tt samplingbincenter}} return lists with the central energies of the bins or sampling points, respectively. For example,
\begin{quote}
{\tt
 \%bc=bincenter() 
}
\end{quote}
A very useful new features is an interpolation function which can directly interpolate a number of
points and evaluate them at a different set of places. For example,
\begin{quote}
{\tt
 \%energ = \{ 4.0,20.0,50.0 \} \\
 \%effs = \{ 0.0,1.0,1.0 \} \\
 \%ires = interpolation(\%energ,\%effs,1,\%bc) 
}
\end{quote}
\index{aedl}{interpolation@{\tt interpolation}}
interpolates the points with x-values {\tt \%energ} and y-values {\tt \%effs} with the interpolation order one (linear interpolation, third parameter) and evaluates the interpolation result at the bin centers obtained above, \ie, it returns a list of the y-values at the places specified by the last parameter. The only allowed interpolation orders are ``1'' (linear) and ``2'' (cubic splines). This example creates
a neutrino factory energy threshold function linearly climbing from 0 to 1 between 4~GeV and 20~GeV. It can be directly used in a channel definition, \eg, 
\begin{quote}
{\tt
 channel(\#nu\_mu\_appearance)< \\
\hspace*{0.5cm}	@channel = \#mu\_plus: +: electron: muon: \#CC: \#MINOS\ \\
\hspace*{0.5cm}	@post\_smearing\_efficiencies = copy(\%ires) \\
>
}
\end{quote}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Experiment definition with \AEDL }

@@

As we have discussed before, \GLOBES\ can only deal with point sources. Thus,  it is not possible to study effects of the finite size of the neutrino production region, such as in the sun or in reactor experiments with many
neutrino sources (\eg, KamLAND). Therefore, a neutrino source in \GLOBES\ can, in general, be characterized by the flux spectrum for each neutrino flavor, the CP sign (neutrinos or antineutrinos), and the total luminosity
of the source.

Before we come to the definition of the source properties, let us discuss
the total integrated luminosity of the experiment. In \GLOBES , the total number of events is in general proportional to the product of
\begin{equation}
N \times \mathrm{Fid.~det.~mass}\,\left[\mathrm{kt/t}\right]\times 
\mathrm{Running~time} \,\left[\mathrm{yr}\right]\times\left\{ \begin{array}{c}
\mathrm{Source~power}\,\left[\mathrm{MW/GW}\right]\\
\mathrm{Useful~parent~decays}\,\left[\mathrm{yr}^{-1}\right]
\end{array}\right.
\end{equation}
with $N$ being a normalization constant.
Thus, the source power corresponds to either the amount of energy produced per time frame in the target (such as for nuclear reactors or sources based on pion decay), or the useful parent particle decays per time frame (neutrino factories, beta beams). In addition, the definition of the source power makes only sense together with the flux normalization, the running time, and the fiducial detector mass in order to define the total integrated luminosity. Therefore, one can, in principle, use arbitrary units for these components as long as their product gives the wanted neutrino flux. However, it is
recommended to use normalizations such that the source power units are $\mathrm{MW}$ for a proton-based beam, and $\mathrm{GW}_\mathrm{thermal}$ for a reactor experiment. Correspondingly, the detector mass units should be kilotons for a proton-based beam, and tons for a reactor experiment. In any case it is a good
idea to document the choices made by the user by corresponding comments
in \AEDL. For more details on the luminosity implementation, see \App~\ref{app:flux}.

@@

\index{aedl}{nuflux@{\tt nuflux}!time@{\tt "@time}}
There are two principal ways to initialize a neutrino flux: 
Either one can use 
a built-in source, or one can provide a file. In both cases,
a flux is defined by the environment {\tt nuflux}, such as
\index{aedl}{nuflux@{\tt nuflux}}
\begin{quote}
  {\tt nuflux(\#name)<\\
\tb $\ldots$\\
\tb @time = 8.0 \\
>}
\end{quote}
with a running time of $8$ years. Note that the running time is used within
the {\tt nuflux} environment. This feature can be used to load the neutrino
and antineutrino fluxes in an accelerator experiment separately, in order
to combine them with different running times in the respective operation modes.
The name of the flux {\tt \#name} will later be referred to in the channel definitions. Note that \GLOBES\ versions older than 3.0 use the still supported {\tt flux} environment, which is different from {\tt nuflux} by an undocumented normalization factor 5.2 for user-defined fluxes. This difference is explained in \App~\ref{app:flux}.

@@

Furthermore, the matter density profile along the baseline
has to be specified. The simplest profile is a constant matter density equal to the average matter density from the PREM~\cite{Dziewonski:1981xy,Stacey} onion shell model of the earth:\index{norm}{PREM| \see{Matter density}}
\index{norm}{Matter density!of the earth}
\begin{quote}
{\tt \$profiletype=1 }
\end{quote}
%
If you are using this option please cite \Refs~\cite{Dziewonski:1981xy,Stacey}\index{norm}{Referencing!matter profile data}.

@@

Cross sections\index{norm}{Cross section} will later be used as part of the 
channel definition (see \Sec~\ref{sec:channel}). Similar to the source 
fluxes, they are provided by the user as a data file:
\index{aedl}{cross@{\tt cross}}
\index{aedl}{cross@{\tt cross}!cross file@{\tt  "@cross\_file}}
\begin{quote}
{\tt cross(\#name)<}\\
{\tt \tb @cross\_file ="user\_file\_1.dat"}\\
{\tt >}
\end{quote}  
This cross section can later be referred to by {\tt \#name}.

Cross sections in \GLOBES\ are given as total cross section divided by energy:
\begin{equation}
\hat\sigma(E)=\sigma(E)/E\,\left[ 10^{-38}\,
\frac{\mathrm{cm}^2}{\mathrm{GeV}} \right]
\end{equation}
The software assumes that the cross section files are text files with 
seven columns and $1001$ lines of the form
\index{norm}{Cross section!file}
\index{aedl}{cross@{\tt cross}!cross file@{\tt  "@cross\_file}}
\begin{quotation}
$\mathrm{log}_{10} E\quad
\hat\sigma_{\nu_e}\quad
\hat\sigma_{\nu_\mu}\quad
\hat\sigma_{\nu_\tau}\quad
\hat\sigma_{\bar\nu_e}\quad
\hat\sigma_{\bar\nu_\mu}\quad
\hat\sigma_{\bar\nu_\tau}$
\end{quotation}
Here the logarithms of the energy values have to be equidistant. For 
arbitrary energies, linear interpolation is used. If the energy leaves the
range of values given in the file, $0.0$ will be assumed. In general, it is advisable to provide the cross sections in the range between {\tt \$sampling\_min} and {\tt \$sampling\_max} (\cf, \Sec~\ref{sec:energy}).
Cross sections for unused neutrino flavors have to be filled with zeros, and
can not just be omitted.

@@

In the following subsections we will explain the energy resolution function.
All energy resolution functions are defined within an {\tt energy} environment and can be referred to by {\tt \#name}:
\begin{quote}
  {\tt energy(\#name)<\\
\tb $\ldots$\\
>}
\end{quote}
The individual parameters of the environment will be defined below and depend on the algorithm used.

@@

\subsection{Low-pass filter$^*$}

@@

\subsection{Manual energy smearing$^*$}

@@

\subsection*{\AEDL\ external variable substitution$^*$}
\index{norm}{globes@{\tt globes}!variable substitution}
